### 2020-03-23 korea.tex
tweaks

\begin{flushleft}{
	\Large
	\textbf\newline{
		Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1,*}
Kaiyuan Sun\textsuperscript{2}
C\'ecile Viboud\textsuperscript{2}
Jonathan Dushoff\textsuperscript{3,4,5}
Bryan T.\ Grenfell\textsuperscript{1,2,6}
\\
\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\textbf{2} Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
\\
\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{5} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{6} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
\\
\bigskip

*Corresponding author: swp2@princeton.edu
\end{flushleft}

@@

Since its first appearance in Wuhan, China, in December 2019 \citep{pneumonia}, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
The epidemic then took off in mid-February when COVID-19 began to spread within a church in the city of Daegu --- as of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church \citep{kcdc}.
While Korea's success in suppressing the epidemic has been widely attributed to intensive testing \citep{fergusonimpact, science}, other factors, such as social distancing, are also likely to have played important roles.
Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by using metro traffic data to compare epidemics in two geographically separated major cities in South Korea.

@@

\begin{figure}
\includegraphics[width=\textwidth]{figure_report_delay.pdf}
\caption{
\textbf{Changes in the number of tests and delay distributions over time.}
}
Vertical lines indicate the date on which testing criteria expanded.
Boxplots (C--D) represent the observed delays.
Black lines and gray ribbons represent the median estimates of the mean delays and their associated 95\% credible intervals.
\label{fig:dist}
\end{figure}

@@

We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
We combined the estimated posterior distribution of delay distributions with previously estimated incubation period distribution (Table 1) to obtain posterior samples for date of infection for each confirmed case.
Likewise, we estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model, while accounting for right-censoring (\fref{dist}D),
to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
Details are provided in the Supplementary Materials.

@@

\fref{eff} compares the estimates of $\mathcal R_t$ in Daegu and Seoul.
In Daegu, estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}A).
Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B);
our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).

>>>>

\begin{flushleft}{
	\Large
	\textbf\newline{
		Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1,*}
Kaiyuan Sun\textsuperscript{2}
Jonathan Dushoff\textsuperscript{3,4,5,$\dagger$}
Bryan T.\ Grenfell\textsuperscript{1,2,6,$\dagger$}
C\'ecile Viboud\textsuperscript{2,$\dagger$}
\\
\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\textbf{2} Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
\\
\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{5} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{6} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
\\
\bigskip

$\dagger$These authors contributed equally\\
*Corresponding author: swp2@princeton.edu
\end{flushleft}

@@

Since its first appearance in Wuhan, China, in December 2019 \citep{pneumonia}, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
The epidemic then took off in mid-February when COVID-19 began to spread within a church in the city of Daegu.
The chains of transmission that began from this cluster distinguish the epidemic in South Korea from that in any other countries:
As of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church and 27\% were in their 20's \citep{kcdc}.
Nonetheless, South Korea's intensive testing using novel tracing techniques allowed it to rapidly identify cases and reduce the spread \citep{fergusonimpact, tracingkorea, science}.
Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by using metro traffic data to compare epidemics in two geographically separated major cities in South Korea.

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_report_delay.pdf}
\caption{
\textbf{Changes in the number of tests and delay distributions over time.}
}
Vertical lines indicate the date on which testing criteria expanded.
Boxplots (C--D) represent the observed delays.
Black lines and gray ribbons represent the median estimates of the mean delays and their associated 95\% credible intervals.
\label{fig:dist}
\end{figure}

@@

We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with the \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
We combined the estimated posterior distribution of delay distributions with previously estimated incubation period distribution (Table 1) to obtain posterior samples for date of infection for each confirmed case.
Likewise, we estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model, while accounting for right-censoring (\fref{dist}D),
to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
We note that changes in the mean decrease delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the increase in the intensity of testing.
Details are provided in the Supplementary Materials.

@@

\fref{eff} compares the estimates of $\mathcal R_t$ in Daegu and Seoul.
In Daegu, estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}A).
Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
The initial decrease in $\mathcal R_t$ may reflect the possibility that a large proportion of church-related transmission could have occurred much earlier; according to the KCDC press release, the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.

On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B);
our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.49$; 95\% CI: 0.20--0.70), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$.
Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).
